---
title: "Practical One"
---

### Finding all row numbers of "airquality" with missing values

```{r echo=-2}
rowsNA <-which((is.na(airquality))) 
rowsNA
```

### Mean of temperature

```{r echo=-2}
tempmean <- mean(airquality$Temp,na.rm=TRUE)
tempmean
```

### Standard deviation of temperature

```{r echo=-2}
tempsd <- sd(airquality$Temp,na.rm=TRUE) 
tempsd
```

### Minimum of temperature

```{r echo=-2}
tempmin <- min(airquality$Temp,na.rm=TRUE)  
tempmin
```

### Maximum of temperature

```{r echo=-2}
tempmax <- max(airquality$Temp,na.rm=TRUE) 
tempmax
```

### Mean of Ozone

```{r echo=-2}
ozonemean <- mean(airquality$Ozone,na.rm=TRUE) 
ozonemean
```

### Standard deviation of Ozone

```{r echo=-2}
ozonesd <- sd(airquality$Ozone,na.rm=TRUE)
ozonesd
```

### Minimum of Ozone

```{r echo=-2}
ozonemin <- min(airquality$Ozone,na.rm=TRUE) 
ozonemin
```

### Maximum of Ozone

```{r echo=-2}
ozonemax <- max(airquality$Ozone,na.rm=TRUE) 
ozonemax
```

### Parameter estimates using matrix calculations

```{r}
yvalues <- cars$dist 
xvalues <- cars$speed 

f <- function(xvalues,yvalues){
 
  y <- as.matrix(yvalues)
  x <- cbind(1,as.matrix(xvalues))
  beta <- solve(t(x)%*%x)%*%t(x)%*%y
  n <- nrow(y)
  k <- ncol(x)
  residual <- (y-x%*%beta)
  fraction <- 1/(n-k)
  rss <-sum(residual^2)
  s <- sqrt(rss*fraction)
  varres <- s^2*solve(t(x)%*%x)
  se <- sqrt(diag(varres))
  return(list("coefficents"=beta,"std.error"=se))
}
modelcalc <- f(xvalues,yvalues)
modelcalc 

```

### Fitting a linear model using lm()

```{r}
modelfit <- lm(dist~speed,data=cars) 
summary(modelfit) 
```

### Comparing coefficients

```{r echo=c(1,4)}
lmcoe <- coef(modelfit)
lmcoe
cat("\n")
calcoe <-modelcalc$coefficents
calcoe 
cat("\n")
```

### Comparing standard errors

```{r echo=c(1,4)}
lmse <- summary(modelfit)$coefficients[, 2]
lmse
cat("\n")
calstd <- modelcalc$std.error
calstd
cat("\n")
```
