---
title: "Practical 3"
editor: visual
---

### Setup

```{r}
install.packages("nycflights13")
library(dplyr)
library(tidyr)
```

### Displaying data in alternative form

```{r}
data(flights,package="nycflights13")
tibble(flights)
```

### Rewriting provided code

```{r}
flights|>filter(month==1)|>group_by(carrier)|>summarise(meandist =mean(distance),stddist=sd(distance))|>arrange(meandist)

```

### Explaining Standard deviation with demonstration 

```{r}
flights|>filter(carrier=="OO") 

flights|>filter(carrier=="HA") |>pull(distance)
```

The standard deviation is NA as there is only one value.

The standard deviation is 0 as all the values for a specific carrier are the same .An example would be HA the distance is 4983 for all therefore the standard deviation value is 0.\

###  Constructing a dataframe  where the carriers are along the columns, and the rows are the average departure delay (`dep_delay`) flown by each carrier (`carrier`) in each month

```{r}
new_df <- flights |> group_by(month,carrier) |> summarise(meandelay = mean(dep_delay,na.rm=TRUE)) |>pivot_wider(names_from=carrier,values_from = meandelay)
new_df
```

### Proportion of flights that were delayed but arrived on time

```{r}
filteredfli <-flights |>filter(dep_delay>0&arr_delay<=0) 
filterlen <-nrow(filteredfli)
originallen<-nrow(flights)
prop <- filterlen/originallen
prop
```

### Routes that more than one airline flies

```{r}
flights |> select(origin,dest,carrier)|>distinct() |>group_by(origin,dest) |> summarise(countc=n()) |>filter(countc>1)
```

###  Average arrival delay for each airline

```{r}
meancalc<-flights|> select(origin,dest,carrier,arr_delay) |>group_by(origin,dest,carrier) |> summarise(meanarr = mean(arr_delay,na.rm=TRUE)) |> left_join(airlines,by= "carrier")
meancalc
```

###  Worst and best average arrival delay 

```{r}
maxdelay <- max(meancalc$meanarr, na.rm = TRUE)
maxname <- meancalc |> filter(meanarr==maxdelay) 
maxname

mindelay <- min(meancalc$meanarr, na.rm = TRUE)
minname <- meancalc |> filter(meanarr==mindelay)
minname
```

###   Reason for difference

\
A possible reason could be too many flights departing early and arriving late

### Indentifying inconsistency 

```{r}
UtilsDataRSV::view_cols(dataset)
```
